---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abctools

<!-- badges: start -->
<!-- badges: end -->

`abctools` is a collection of functions useful across different ABC analyses and
projects.

## Installation

You can install the development version of abctools from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ABC-Kenya-Study/abctools")
```

## Getting started

`abctools::` has a friendly wrapper for collecting ABC data from REDCap:

```{r get.abc.data}
library(abctools)

# path to .R script containing ABC API token, e.g., "abc_token.R"
tp <- file.path(Sys.getenv("R_USER"), "R", "access_tokens", "abc_token.R")

# collect REDCap data
abc_data <- get_redcap_data(token_path = tp)
```


## Helper functions

Once you have ABC data, `abctools::` also has helper functions for wrangling
tasks. For instance, `drop_vars_all_na()` deletes columns that are all NA:

```{r wrangle.abc, warning=FALSE, message=FALSE}

library(dplyr)
library(stringr)

ncol(abc_data)

enroll <- abc_data |>
  filter(str_detect(redcap_event_name, 'Enrollment')) |>
  drop_vars_all_na()

enroll

ncol(enroll)
```

Enjoy!
