---
output: github_document
editor_options: 
  markdown: 
    wrap: 72
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# abctools

<!-- badges: start -->

<!-- badges: end -->

`abctools` is a collection of functions useful across different ABC
analyses and projects.

## Installation

You can install the development version of abctools from
[GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ABC-Kenya-Study/abctools")
```

## Getting started

`abctools::` has a friendly wrapper for collecting ABC data from REDCap:

```{r get.abc.data}
library(abctools)

# path to .R script containing ABC API token, e.g., "abc_token.R"
tp <- file.path(Sys.getenv("R_USER"), "R", "access_tokens", "abc_token.R")

# collect REDCap data
abc_data <- get_redcap_data(token_path = tp)
```

## Helper functions

Once you have ABC data, `abctools::` also has helper functions for
wrangling tasks. For instance, `drop_vars_all_na()` deletes columns that
are all NA:

```{r wrangle.abc.data, warning=FALSE, message=FALSE}

library(dplyr)
library(stringr)

# original number of columns in data from REDCap
ncol(abc_data)

# subset to enrollment rows, then drop columns with all NAs
enroll <- abc_data |>
  filter(str_detect(redcap_event_name, 'Enrollment')) |>
  drop_vars_all_na()

enroll

# number of columns in 'enroll'
ncol(enroll)
```

## Derived variables

Functions that create derived variables are also included in `abctools`.
Derived variables are have the `calc_` prefix. Revisit this page to
check for more functions as they are added.

```{r derived.variables}

# assess participants' exposure to environmental tobacco smoke (ETS)
ets <- enroll |> 
  # subset to relevant variables (optional)
  select(cp_ptid, lbt_closetosmokers, en_peoplesmoking, en_ind_cig, en_ind_marig, 
         en_wrk_cig, en_wrk_marij) |> 
  ets_exposed(
    freq_factor = list(no = c('Not at all', 'Rarely'), 
                       yes = c('Some days', 'Most days', 'Daily'))
    )

# tabulate ETS status
table(ets$calc_ets_exposed)
```

```{r plot.ets, echo=FALSE, warning=FALSE}

library(ggplot2)

# calculate percentages
ets |> 
  group_by(calc_ets_exposed) %>% 
  summarise(count = n()) %>% 
  mutate(perc_ets_exposed = 100*count/sum(count)) |> 
  ungroup() |>

# plot ETS status
ggplot(aes(x = calc_ets_exposed, y = perc_ets_exposed, fill = calc_ets_exposed)) + 
  geom_bar(stat = "identity") +
  labs(fill = "ETS exposed", y = "Percentage of Participants (%)" ) +
  scale_fill_brewer(palette = "Set1") + 
  scale_y_continuous(expand = expansion(mult = c(0, .025)) ) +
  theme_bw() + 
  theme(axis.title.x = element_blank(), legend.position = "bottom") 
```

## Output data

`abctools::` also has a straightforward workflow to prepare a flat, wide
dataset with one participant per row. Data cleaning functions and
functions that create derived variables are designed to be applied to
the data collected from REDCap, beforehand.

```{r wide.data}

wide_data <- abc_data |> 
  
  # subset to moms
  filter(str_detect(redcap_event_name, 'Mothers')) |> 
  
  # apply data cleaning functions from `abctools::`
  fix_ptid() |>
  char_to_numeric() |>
  remove_returns() |>
  
  # add derived variables from `abctools::`
  active_smoker(
    freq_factor = list(no = c('Not at all', 'Rarely'), 
                       yes = c('Some days', 'Most days', 'Daily')) 
    ) |> 
  
  # add REDCap event name (CRF) abbreviations
  abb_redcap_events(long_short = 'short') |> 
  
  # make wide data, one row per participant 
  # (note: `wide_data()` needs variable "redcap_event" from `abb_redcap_events()`)
  wide_data()

# show
wide_data

# number of rows and columns of wide dataset
dim(wide_data)
```

Enjoy!
